RadLib: a radiative property model library for combustion CFD

### 2021-05-04 submissions/CPC_response_to_reviewers.md

    3. These kinds of statements persist throughout the second paragraph as well. There are limitations in RadLib such as lack of scattering treatment for particle-laden flows, inability to include more species, lack of any line-by-line capabilities, etc. The library checks a few useful boxes, but the verbiage here seems to take more credit than the authors may be due.

    **The authors have a number of concerns about the reviewer's claims here that they would like to address. First, note that the subject of the first sentence of Section 5 is "radiative heat transfer models", which does not refer specifically to either radiation property models or RTE evaluation methods, but the treatment of computational radiative heat transfer as a whole. This generalization holds for the entire first paragraph of Section 5. Additionally, at no point in the paper do the authors claim that property evaluation is more complex or computationally expensive than the RTE solution, only that it "can be a difficult and time-consuming portion of radiation calculations" (pg. 2). Likewise, the paper contains no claims about the complexity of radiation property models in relation to other aspects of combustion simulations. (On this topic, however, the authors would like to point out that these other aspects of combustion simulations are typically well-characterized in both commercial and open-source CFD codes and do not require very much user interaction, whereas radiation modeling is often absent or highly simplified.) The second paragraph of Section 5 is a brief discussion on how RadLib can benefit researchers that draws from the authors' experiences. It does not claim that RadLib will solve any of the challenges of radiation modeling, only that it can "ease . . . obstacles", which the authors believe it can do.**

    **All of that said, however, the fact that the reviewer made these assumptions and pointed out these implications indicates that the paper's language is, in fact, too subjective. Even though none of these statements were explicitly made, they must be implied for the reviewer to come to these conclusions. As such, revisions have been made throughout the paper to reduce subjective language, support generalized statements with additional explanation and/or references, and include due discussion of RadLib's limitations. Section 5 in particular has undergone heavy revision.**

@@

2. make_examples.sh:
    1.	will only work with a g++ compiler in the user’s path.
    1.	Is referred to as build_examples.sh in the README.md file
    1.	is redundant with the cmake build system and isn’t portable. It should probably be removed.

	**The ```make_examples.sh``` file has been removed.**

3. It appears that there are no tests associated with the library. This is a bit surprising - I would expect regression test coverage on the basic API functionality.

	**The authors acknowledge the lack of tests associated with the library and note that basic test coverage is currently being developed.**

@@

5. CMake build system:
    1.	Consider installing a “RadLib.cmake” file for downstream usage by CMake-based projects. This helps downstream build systems configure for RadLib usage (setting include paths, etc.).

		**Installation now includes a ```radlib.cmake``` file, located by default in the ```installed/cmake/radlib``` directory, for downstream usage.**

@@

    3.	Using GLOB to install the headers is not the appropriate CMake approach.

		**Build updated accordingly; GLOB no longer used to install header files.**

    5.	the data files are not copied into the examples build directory, meaning that the example executables do not run.

		**The authors have experienced no issues running the example executables from either the ```build/examples``` directory or directly from the ```examples``` directory. More information is required to address this issue.**

@@

    7.	The run-examples.sh script is not helpful as it is not installed into the build directory where the executables are produced and doesn’t have path information to actually run the executables.

		**Executables are built within the example folders in addition to the build directories, so ```run_examples.sh``` runs without any issues.**

@@

5. It appears that line-by-line data from HITEMP/HITRAN files cannot be loaded into RadLib. That’s an unfortunate limitation, but not a deal-breaker.

	**At present, RadLib's models cannot accommodate such data input, but this is a limitation of the models themselves, not the software. The appropriate line-by-line data is included within the RadLib package, which the authors consider a convenience rather than a hindrance. The ability to load database files directly into RadLib could be useful to researchers looking to extend RadLib's model library to include modern spectral models, however. The authors appreciate the suggestion and will consider it for future development.**

6. It looks like the software hard-codes Plank mean absorption coefficients as (fourth-order?) polynomial fits? Same for WSGG. There is no discussion of why this was chosen, how the coefficients were determined, or a characterization of its accuracy. Do you observe any problematic behavior with the fourth order polynomials over temperatures from 300 to 3000 K?  Are these interpolants or regression of data?

	**This discussion was purposely omitted because it does not align with the goal of the paper, which is to present RadLib as a software tool for calculating radiation properties, not to evaluate the included models. All of the reviewer's questions are addressed in detail in the referenced literature (Planck Mean coefficients in [3,5] and WSGG in [10,11]). The authors believe that this is an issue of scope rather than missing information but are willing to extend the discussion at the editor's discretion.**

@@

8. RadLib only considers CO2 and H2O for WSGG, which is fairly limited. Even for Plank mean, only H2O, CO2, CO and CH4 are considered. NO and OH would be nice to include. It appears that including additional species would be a non-trivial undertaking.

	**All WSGG correlations consider only CO2-H2O mixtures with the exception of specialized subset of WSGG models that consider CO2-H2O-soot mixtures. This WSGG model was chosen for RadLib due to its popularity, relative simplicity, and good performance. RadLib implements the WSGG model presented by Bordbar et al. in references [10] and [11], which was developed specifically for use in oxy-fired combustion simulations. CO2 and H2O are the major products of all combustion processes, and CO is a major product of rich combustion. NO and OH are minor species that, while important to other aspects of combustion processes, do not typically participate significantly in radiative heat transfer processes. For more information, please refer to Section 20.6 of reference [2] and/or Section 3.4.1 of *Radiative Heat Transfer in Turbulent Combustion Systems* by Modest and Haworth.**

	**It should also be noted that the Planck Mean absorption coefficients model as implemented in RadLib refers specifically to the radiation model in [3,5], which is one of the most common radiation models applied to turbulent combustion simulations. This is why Section 2.1 explicitly references the TNF radiation model and its calculated Planck Mean absorption coefficients. As implemented, the model is not intended to be general or extensible to other gas species and was never presented as such. Adding additional species would indeed be a non-trivial undertaking, but that is neither the goal nor purpose of RadLib as a software package, which presents and validates models rather than evaluating them.**

	**Major revisions have been made to the paper text to clarify points like these for readers who do not specialize in combustion processes or modeling and provide additional information about why these models were chosen for implementation in RadLib.**

@@

	**Build process has been updated so that documentation can be generated by the system target or manually.**

## Reviewer 2

A radiation spectral property library is developed in this study using C++. A Planck-mean absorption coefficient based model, a WSGG model and a RCSLM model from the literature are implemented  in the package. Verification results are presented and discussed. Although the authors addressed an important issue of radiation modeling, the contribution is marginal in the reviewer's opinion. No new models/methods are proposed in this study. As the authors correctly pointed out, the solver for radiative transfer equation is equally important, which is not included in the library. The coupling of a spectral model to a specific RTE solver, can have different levels of challenges depending on the combination. The design of the library doesn't really address any of the RTE-related issues, nor provide any new solutions to simplify the coupling of spectral models to RTEs. If the authors can demonstrate that the library can be coupled with popular CFD software and popular RTE solvers such as P1 and DOM, it would further improve the quality of this study. A few other concerns are listed as follows:

**The authors would like to thank the reviewer for their comments and suggestions. We have already begun working on coupling RadLib to CFD software (that includes a DOM solver for the RTE solution, among other features) in order to demonstrate RadLib's capabilities.**

@@

2. Page 10, missing information on [INSERT RUNTIME COMPARISONS HERE].

	**This was a placeholder statement that should not have been included in the document. It has been removed.**

3. In Figure 2, the RCSLM seems to consistently perform better than WSGG, except in Example S2. In addition, WSGG performs worse than PM in Example S4. Can the authors explain why?

	**The PM and WSGG models are primarily empirical and contain very little representative physics, which could explain their unexpected behavior in Examples 2 and 4 but also makes it very difficult to find or treat sources of error. Any model, but especially empirical models, can behave unexpectedly when applied to conditions or parameters outside of those for which it was developed. The example cases are meant to be illusrative of the software's capabilities, not an evaluation of the models themselves, which is outside the scope of this paper. The reviewer's request for an explanation, however, is valid. The discussion in Section 5 has been expanded and revised to address these specific discrepancies. Additional information about the specific limitations, strengths, and weaknesses associated with each model can be found in Section 2.**

@@

	**The authors purposely chose the TNF workshop data for calculating Planck Mean absorption coefficients because it is one of the most common and widely used radiation models applied to turbulent combustion simulations, despite its age and limitations. Section 2.1 has been revised to include a more explicit dicsussion of why the TNF radiation model was chosen and concerns related to its relative age and limited applicability. Section 4 includes example cases that clearly demonstrate the deficiencies of the PM model compared to the other models included with RadLib, and some additional discussion is available there.**

5) On page 13, the authors present a paragraph describing Figs. 2 and 3 - the comparison of various radiation property models present in RadLib in one-dimensional configurations. The comparisons are presented simply as observations without any explanation or reasoning why the models behave the way as seen in Fig. 2 and 3. If I understand correctly, the focus of the work presented here is not the comparison of the accuracy of the three models but to showcase the library. Therefore, these comparisons - specifically without any explanation of the results - may not be essential to the current manuscript, and can be shortened. These results do serve as descriptions of example cases provided in RadLib. On the same page, the authors say that "While ommitted [should be omitted] here for brevity, the implemented WSGG and RCSLW models give essentially identical results to those presented in [11, 14] such that these examples also serve as a validation of the implementation of the
models." The authors should avoid qualitative description such as "essentially identical" in the context of validation and present a quantitative (e.g., % error) metric wherever possible. A strong validation is important for programs like this.

	**While it is true that the purpose of this paper is to present the RadLib library rather than evaluate or compare the models, the authors believe that example case results in Figures 2 and 3 are important to the illustration of RadLib's capabilities. The reviewer's point about presentation without explanation, however, is valid; to that end, the discussion in Section 5 has been expanded to acknowledge, if not fully explain or provide cause for, various features and descrepancies in the example case results. Due to the empirical nature of the PM and WSGG models, it is very difficult to identify, explain, or rectify unexpected behavior.**

@@

	**In regards to validation, the authors are unclear as to what the reviewer expects to see. RadLib's model implementations give results that are, in fact, essentially identical to those presented in the literature to within any visual margin. We do not have access to the data used to generate the plots in the literature, without which any quantitative measure of error is meaningless. We do, however, compare model results to the line-by-line data, which amounts to a direct solution of the RTE and is also identical to the line-by-line data in the literature (as expected). The authors believe that this serves as adequate validation of RadLib's model implementations given the lack of originial data for comparison.**

6) The computational cost tables for two cases show a significantly different trend, which the authors attribute to "additional overhead required in performing the calculations." Can the authors be more specific as to what these overheads may be? They say that they have neglected the initialization and input/output costs. Also, do these CPU times presented include the cost for the ray-tracing solver as well? Are the grids kept the same for all configurations?

	**Section 4.1 already includes details of the computational cost calculations presented in Figure 4. The left hand plot (a) refers to one million evaluations of the absorption coefficients directly, while the right hand plot (b) refers to the cost to evaluate Example S3, which, like all of the example cases described in the first paragraph of Section 4, includes the ray-tracing solver. The quoted sentence has been revised to specify that the ray-tracing solver is primarily responsible for the additional overhead. Grids are identical for each configuration. Note that the RadLib library does not generate or manipulate grids as it is not involved in the RTE solution. The ray-tracing solver used in the example cases is provided for convenience and demonstration purposes.**

	**Note that Section 4.1 has since been moved to Section 5. It and the appropriate figure captions have also been revised for clearer communication of the points above.**

@@

2) In the program summary the description of the nature of the program is somewhat confusing from a grammatical construction standpoint. Particularly, the following part, "This presents a problem for practitioners who wish to use/implement radiation models for which [refers to what?] such models consitute [constitute?] only a small but important part of a larger simulation.  Turbulent combustion simulations are one such example [example of what?].  Often, rudimentary assumptions [in what?] are made and this can negatively impact results."

	**The program summary has been rewritten for clarity.**

3) In Page 2, "In some cases, radiative gains or losses are small compared to other energy sources or heat transfer modes and radiation can be safely neglected, ... difficult to simulate accurately without a robust radiation model." Please provide some references to support these arguments.

	**References have been added throughout the text to bolster weakly supported arguments. Due to unrelated revisions, the quoted statement and related discussion is now located in Section 2.1 since it is most closely related to the Planck Mean model and its limitations.**

@@

6) In lines 48-52 on Page 4, the authors say that the absorption coefficients are calculated using some correlations after some curve-fitting to spectroscopic databases. I am not sure what the authors refer to as correlations. There may be some correlations involved in WSGG models, but it is not clear to me where are correlations involved in RCSLW.

	**The RCSLW model's key tool is the absorption line blackbody distribution function (ALBDF), which is calculated using sppectral data available from databases like HITEMP. This is discussed briefly at the end of Section 2.3 on pages 8-9. For more information on how spectral data is used to calculate the ALBDFs, please see reference [22].**

7) On page 7, the authors mention that the PM model can be "reasonably accurate" in some cases and go on to say that, "The Planck Mean model is most appropriate under optically thin conditions  with  relatively  low  radiative  transfer  relative  to  other  heat  sources such  as  reactive  heat  release  rates." The Planck Mean model will be appropriate in optically thin cases as there is very little reabsorption. This may or may not be connected with "relatively  low  radiative  transfer  relative  to  other  heat  sources." In fact, if the radiative heat transfer is much lower than chemical heat release, neglecting radiation altogether may also be reasonable! Although, the definition of "relatively low" is subjective.

	**The reviewer's statements regarding PM absorption coefficients are generally true. In combustion simulations, low radiative heat transfer relative to other heat sources, including chemical heat release and convective heat transfer, usually does, in fact, indicate very low radiative reabsorption rates. Sectiom 2.1 and related discussion elsewhere has been revised to clarify this connection. The quoted sentence now reads as follows: "The Planck Mean model is most appropriate under optically thin conditions in which low radiative transfer relative to other heat sources, including but not limited to chemical heat release and buoyancy-driven convection in combustion systems, indicates that very little radiative reabsorption is occuring."**

### 2021-07-19 submissions/CPC_response_to_reviewers.md

## Reviewer 1

This manuscript introduces “RadLib,” a C++ library to calculate absorption coefficients for use in radiative heat transfer calculations. The library has reasonable value to the research community, despite a few limitations in its scope and design. The library is self-contained (doesn’t have any external dependencies), which is convenient.

I have a number of comments on the manuscript as well as the software itself which the authors should consider to improve the manuscript and software.

**The authors would like to thank the reviewer for their time and careful review.**

### 2021-07-19 manuscript/manuscript.tex
Editing the software section: fortran, get_k_a_onebranch

RadLib is an object-oriented C++ class library that includes both C++ and Python interfaces. Upon initial download, the RadLib package contains five subdirectories: \texttt{src} contains the RadLib source code; \texttt{build} contains installation files; \texttt{examples} contains instructive example cases, including a simple interface and solver for a parallel planes geometry; \texttt{data} contains ALBDF data tables required for the RCSLW model; and \texttt{docs} contains files used to optionally generate code documentation with Doxygen. An additional directory, \texttt{installed}, is created and populated with RadLib's executables and header files during installation if no other installation location is specified.

The \texttt{source} and \texttt{examples} directories are further divided into \texttt{C++} and \texttt{Python} subdirectories to differentiate between interfaces. There are three interface options for using this code: C++, Python, and Cython-wrapped Python. C++ interfaces are located within the \texttt{examples/c++} folder, while Python interfaces are located within the \texttt{examples/python} folder. When running examples with Python, the Cython-wrapped version is the default; to run the regular Python version without the Cython wrapper, edit the comments near the top of the example files (i.e \texttt{ex\textunderscore S1.py}). The C++ interface produces the fastest-running code, followed by the Cython-wrapped Python interface and then the regular Python interface.

RadLib installation is automated with CMake from within the \texttt{build} directory. If the user requires an installation location other than the default \texttt{installed} directory, edit the "Project Options" section of \texttt{build/CMakeLists.txt} file before running CMake. First, configure CMake by running the command \texttt{cmake .} from within the \texttt{build} directory. Upon successful completion, run \texttt{make} to compile the library and then \texttt{make install} to install it. To use RadLib in C++ projects, include the generated C++ library file, located at \texttt{installled/lib/libradlib.a}, and/or header files, located in \texttt{installed/include}, according to the project's build scheme. Installation also generates a relocatable RadLib package, located at \texttt{build/cmake/radlib/radlib.cmake}, for downstream usage. To use RadLib in Python, update the system \texttt{PYTHONPATH} environment variable and import the module, named \texttt{pyrad}, into the project.

The interfaces, solver, and examples included with the library serve to demonstrate its use and validate the implementation and results. Figure \ref{fig:flowchart} illustrates the basic structure use of the RadLib package within a generic example using the provided interfaces. The RadLib library generates absorption coefficients and their weighting factors for use within the appropriate RTE, but does not specify any particular case geometry. After creating a radiation object, users interact with RadLib through the \texttt{get\textunderscore k\textunderscore a()} function, which requires arguments corresponding to the system temperature; system pressure; species mass fractions for H$_2$O, CO$_2$, CO, and CH$_4$; soot volume fraction in units of parts per million (ppm), if applicable; and variables initialized to receive output values for the radiation absorption coefficients and their corresponding weights. For examples that illustrate how a user might interact with RadLib, refer to the \texttt{simple\textunderscore interface.cc} and \texttt{simple\textunderscore interface.py} example files located in the \texttt{examples/c++} and \texttt{examples/python} folders, respectively. For information about adding additional radiative property models to the RadLib library, please refer to the package documentation. 
%
\begin{figure}
	\begin{center}
        \includegraphics[width=\textwidth]{fig_radlib_structure.pdf}
	\end{center}
	\caption{Example workflow diagram. Highlighted areas are part of the RadLib library; other areas represent example infrastructure for using the package.}
\label{fig:flowchart}
\end{figure}
%

>>>>

RadLib is an object-oriented C++ library that includes C++, Python, and Fortran interfaces. Upon initial download, the RadLib package contains four subdirectories: \texttt{src} contains the RadLib source code; \texttt{examples} contains instructive example cases, including a simple interface and solver for a parallel planes geometry; \texttt{data} contains ALBDF data tables required for the RCSLW model; and \texttt{docs} contains files used to optionally generate code documentation with Doxygen. An additional directory, \texttt{installed}, is created and populated with RadLib's executables and header files during installation if no other installation location is specified.
The \texttt{src} and \texttt{examples} directories are further divided into \texttt{c++}, \texttt{python}, and \texttt{fortran} subdirectories to differentiate between interfaces. 
%There are three interface options for using this code: C++, Python, and Cython-wrapped Python. 
%C++ interfaces are located within the \texttt{examples/c++} folder, while Python interfaces are located within the \texttt{examples/python} folder. 
%When running examples with Python, the Cython-wrapped version is the default; to run the regular Python version without the Cython wrapper, edit the comments near the top of the example files (i.e \texttt{ex\textunderscore S1.py}). The C++ interface produces the fastest-running code, followed by the Cython-wrapped Python interface and then the regular Python interface.

RadLib installation is automated with CMake as follows. Inside the downloaded \texttt{radlib} folder, the user creates and enters a \texttt{build} directory, then enters the commands \texttt{cmake ..}, followed by \texttt{make}, followed by \texttt{make install}. Project options are specified by editing the top-level \texttt{CMakeLists.txt} file where the options are clearly marked. These can also be specified at the command line. Options include specifying the system installation location, which defaults to \texttt{radlib/installed}; options to build the Python and Fortran interfaces; options to build examples for C++, Python, and fortran; and an option to build the documentation.

The code installation folder includes directories \texttt{include} containing header files and a fortran module file, \texttt{lib} containing library files for each interface, and \texttt{rcslw_data} containing binary data files used by the RCSLW model. These binary files contain the ALBDF functions for a given species at a given pressure (in atm units), as indicated by the file name. The \texttt{include} directory contains the following interface header and module files:
%
\begin{itemize}
    \item \texttt{rad_planck_mean.h},
    \item \texttt{rad_wsgg.h},
    \item \texttt{rad_rcslw.h},
    \item \texttt{rad_module.mod}.
\end{itemize}
%
The C++ and Fortran libraries in the \texttt{lib} folder are
\begin{itemize}
    \item \texttt{libradlib.a},
    \item \texttt{libradlib_fortran.a},
\end{itemize}
%
This folder also contains the python module.
The code installation additionally generates a relocatable CMake package for downstream usage, located at \texttt{build/cmake/radlib/radlib.cmake}. If the Python interface is built, the location of the installed Python package needs to be added to the \texttt{PYTHONPATH} environment variable. A command to do this is provided at the end of the installation (for example, \texttt{export PYTHONPATH="/home/user/radlib/installed/lib/python3.8/site-packages:$PYTHONPATH"}).

To use RadLib in C++, a user code includes the header file for the desired radiation model, e.g., \texttt{#include "rad_rcslw.h"}. For Fortran, a \texttt{use rad_module} statement provides access to the required interface functions. The Python interface functions are loaded from the \texttt{pyrad} module, as in \texttt{from pyrad import pyrad_rcslw}. Such usage is demonstrated in the code examples, discussed below. Compilation of C++ and Fortran programs requires specification of the header/module and library directory locations and linking against the installed libraries (\texttt{libradlib.a} for C++ and both \texttt{libradlib.a} and \texttt{libradlib_fortran.a} for Fortran).

Figure \ref{fig:flowchart} illustrates the basic structure of using of RadLib in a generic one-dimensional RTE solution. The RadLib library generates absorption coefficients and their weighting factors for use within an appropriate RTE solver. A user first creates a radiation object, and then makes calls to one of two primary functions: \texttt{get\textunderscore k\textunderscore a()}, or \texttt{get\textunderscore k\textunderscore a\textunderscore oneband()}.
These two functions take the arguments for temperature (K), pressure (Pa), soot volume fraction, and mole fractions of \ce{H2O}, \ce{CO2}, \ce{CO}, and \ce{CH4}. The WSGG model does not account for \ce{CO} or \ce{CH4}, and only the PM model includes \ce{CH4}. A uniform interface is facilitated by giving default values of 0.0 for the \ce{CO} and \ce{CH4} mole fraction parameters. The \texttt{get\textunderscore k\textunderscore a()} function computes a vector of absorption coefficients and a vector of weight factors, with elements corresponding to each clear and gray gas considered in the model (with the clear gas coming first). The PM model only considers one gas and the vectors are of unity size. The \texttt{get\textunderscore k\textunderscore a\textunderscore oneband()} computes the absorption coefficient and weight factor (as scalar values) for a single gas (or \emph{band}), whose index (starting at zero for all interfaces) is provided as an additional argument. This option to compute the properties for a single gas/band can be important for efficient implementation in some solvers, e.g., in the Fire Dynamics Simulation (FDS), discussed below \cite{FDS}. The absorption coefficients and weights are returned through the function argument list for the C++ and Fortran interfaces, and as return values for the Python interface. Additional details are included in the program documentation and code listing.
%
\begin{figure}
	\begin{center}
        \includegraphics[width=\textwidth]{fig_radlib_structure.pdf}
	\end{center}
	\caption{Example workflow diagram. Highlighted areas are part of the RadLib library; other areas represent example infrastructure for using the package.}
\label{fig:flowchart}
\end{figure}
%
